% Since \input doesn't like preambles, only uncomment next line if you're editing this TeX, don't forget the \end{document}! 
%\documentclass{article}\input{../MAIN/packages}\input{../MAIN/commands}\begin{document}
\section{Special Functions}  
    Special functions functions that are defined to have established names and properties because of their importance. Simple examples of special functions are the {\bf sin and log} functions. Some of them are unintuitive, and the notation is often weird, but analysis becomes easier if you use them. 
    \subsection{Bessel Functions}
        Bessel functions are defined to be solutions to Bessel's differential equation:
        \begin{gather}
            x^2y''(x)+xy'(x)+\qty(\lambda x^2-n^2)y(x)=0\quad\parametric{ n\in\mathbb{N}\\\lambda>0\\x\geq0}
            \\
            y(x)=c_1J_n(\sqrt{\lambda}x)
        \end{gather}
        Where $J_n$ is a Bessel function of the first kind. 
    \subsection{Dirac Delta Function}
        The Dirac delta function $\delta(x)$ is a generalized distribution defined by Paul Dirac to have the following properties:
        \begin{align*}
            \delta(x)\equiv\parametric{
            \infty & x=0
            \\
            0 & x\neq 0
            \\[0.25 cm]
            \int_{-\infty}^\infty \delta(x) dx &= 1}
        \end{align*}
        The Dirac Delta Function picks out the value of the function $f$ at a single point.
        \begin{align*}
            \intall f(x) \delta(x-a) dx =\intall f(a) \delta(x-a)dx
            &= f(a)
        \end{align*}
        Note that the interval of integration only needs to contain $a$ for this to be true. 
        \subsubsection*{Three-dimensional Dirac Delta Function}
            
            \begin{align*}
                {\delta}^3(r) &= \delta(x) \delta(y) \delta(z)
            \end{align*}
            \begin{align*}
                \intall f(r) {\delta}^3(r-a) dV &= f(a)
            \end{align*}
            A useful version of this, re-casted for use in electrodynamics:
            \begin{align*}
                \div\qty(\frac{\e{r}}{{r}^2}) &= 4 \pi {\delta}^3(\vb{r})
            \end{align*}\newpage
    \subsection{Error Function}
    \subsection{Gamma Function}
    \subsection{Hermite Polynomials}
    \subsection{Laguerre Polynomials}
    \subsection{Legendre Polynomials}
        These are a set of complete and orthogonal polynomials which are handy in electrodynamics and quantum mechanics. 
        %\tikzsetnextfilename{legendre}
        They can be generated by the Rodriguez formula\cite{pinsky_2011}:
        \begin{equation}
            P_n(x)=\oo{2^nn!}\der[n]{x}\qty[\qty(x^2-1)^n]
        \end{equation}
        \begin{figure}[h!]
            \centering
            \begin{tikzpicture}[scale=1.25]
                \begin{axis}[range=-1:1,domain=-1:1, thick, legend pos= outer north east]
                    \addplot[samples=2]{1};
                    \addplot[samples=10, blue]{x};
                    \addplot[samples=20, purple]{(3*x^2-1)/2};
                            \addplot[samples=40, yellow!50!orange]{(5*x^3-3*x)/2};
                            \legend{1, 
                            $x$,
                            $\oo{2}\qty(3x^2-1)$,
                            $\oo{2}\qty(5x^3-3x)$}
                \end{axis}
            \end{tikzpicture}
            \caption{The First four Legendre Polynomials ($n=0,1,2,3$)}
            \label{fig:legendre_graph}
        \end{figure}
%\end{document}