\documentclass{article}
\input{../../commands.tex}
\header{Physics 457}{Homework 1}{Evan Carpenter}
\thetitle{Physics 457 Hw 1}
\date{Due: 01-19-2022}
\author{\EC}
\DeclareMathOperator{\GeV}{GeV}
\begin{document}
\maketitle
\begin{problem}
    Make the following unit conversions, and show all work. 
    Note that this tests how you use natural units where $c = \hbar = k_B = 1$. Thus, there is one 
    fundamental dimension, which we can take to be energy and express it in GeV to various 
    powers.
    \begin{enumerate}[label=\alph*)]
        \item 1 GeV, in joules (energy) 
        \item 1 GeV, in Kelvin (temperature) 
        \item 1 GeV, in kilograms (mass) 
        \item 1 GeV$^{-1}$, in meters (length) 
        \item 1 GeV$^{-1}$, in seconds (time) 
        \item 1 GeV$^{4}$, in kg/m$^3$ (mass density) 
    \end{enumerate}
    \answerline
    \begin{enumerate}[label=\alph*)]
        \item $1~GeV= 1.602\times10^{-10}~J$ 
        \item $k_B=8.617 333 262\times 10^{-5} eV/K=1
            \rightarrow
            8.617 333 262\times 10^{-14} \GeV=1K\rightarrow \coloredanswer{1\GeV=1.16045\times10^{13}K}$
        \item $E=mc^2$ so $E/c^2=m\rightarrow\coloredanswer{1\GeV=1.783 \times10^{-27} kg}$ 
        \item $E=\frac{hc}{\lambda}$ so $1\GeV^{-1}=\frac{\lambda}{hc}\rightarrow \coloredanswer{1\GeV^{-1}=1.973\times10^{-16}m}$
        \item $\hbar=6.58212\times10^{-16}eVs\rightarrow\frac{\hbar}{eV}=6.58212\times10^{-16}s$
        \\
        $\rightarrow$$\frac{1}{eV}=6.58212\times10^{-16}s\rightarrow\coloredanswer{\GeV^{-1}=6.58212\times10^{-25}s}$
        \item Using results from c),d), the answer is:
        \begin{gather*}
            \frac{1\GeV}{1\GeV^{-3}}=\frac{1.783 \times10^{-27} kg}{\qty(1.973\times10^{-16}m)^3}
            \\
            \coloredanswer{1\GeV^{4}=2.322\times10^{20} kg/m^{3}}
        \end{gather*} 
    \end{enumerate}
\end{problem}
\newpage
\begin{problem}
    Unstable particles appear to live longer if moving, therefore can travel a longer distance after creation. Consider the following problems and calculate the flight distances. 
    \begin{enumerate}[label=\alph*.]
        \item Calculate the flight distance of a muon with 100 GeV energy. Note that the muon lifetime (in its rest frame) is $2.2 \mu s$. 
        \item Calculate the flight distance of a B$^{+}$-meson with 20 GeV energy if its lifetime is $1.6\times10^{-12}$ s and its mass is 5.38 GeV. 
        \item Pions are produced in the upper atmosphere when a proton from outer space hits a proton in the atmosphere. The pions then decay into muons:
        \begin{gather*} 
            \pi^{-}\rightarrow \mu^{-}+\overline{\nu}_\mu
            \\
            \pi^{+}\rightarrow \mu^{+}+\nu_\mu
        \end{gather*}
        But the lifetime of the pion (2.6 $\times10^{-8}$ s) is much shorter than that of the muon. If the pion is produced at 800 meters above the ground, can it reach the ground if its speed is 0.998c?
    \end{enumerate}
    \answerline
    \begin{enumerate}[label=\alph*.]
        \item Given the energy and $\Delta t_{lab}$ and knowing 
        \begin{gather*}
            d_{flight}=v_{lab}\Delta t_{lab}=\boxed{v\gamma\Delta t_{\mu}}
            \\
            \gamma=E/m_\mu\approxeq 943.4\implies v\approxeq1(c)
            \\
            \Delta t_\mu=2.2\mu s
            \\
            m_\mu=0.1056\GeV/c^2
        \end{gather*} 
        from lecture, I can calculate:
        \begin{equation*}
            d_{flight}=943.4*3\times10^{8}\frac{m}{s}*2.2\times10^{-6}s=622,644m
            \rightarrow
            \coloredanswer{d_{flight}=622~\text{km}}
        \end{equation*}
        \item $\gamma=\frac{E}{m_{B^+}}=3.71747\rightarrow v\approxeq0.96314c$
        \begin{equation*}
            \coloredanswer{d_{flight}=0.001717~\text{m}}
        \end{equation*}
        \item $\displaystyle\gamma=\qty(1-\qty(\frac{v}{c})^2)^{-1/2}\approxeq 15.819$
        \begin{equation*}
            d_{flight}=0.998c\times15.819\times2.6\times10^{-8}s=\coloredanswer{123.056~\text{m}}
        \end{equation*}
    \end{enumerate}
\end{problem}
\newpage
\begin{problem}
    Antiprotons were first created at Lawrence Berkeley National Lab (LBL) in 1955 by a proton beam hitting a proton target with the following reaction:  
    $$p+p\rightarrow 3p+\overline{p}$$  What is the minimum total energy $E$ of the proton beam to allow this reaction? Please give your answer in unit of proton mass $m_p$. (Hint: using center-of-mass energy $E_{CM}$ conservation, and assume the final particles are produced at rest.)
    \answerline
\end{problem}
\end{document}