\documentclass{article}
\input{../../commands}
\thetitle{Statistical Mechanics}[\tt Physics 406 at \rm\small \UM]
\date{Winter Semester 2022}
\author{\EC}
    \def \pa {particle\ }
    \def \pas {particles\ }
    \def \ste {state\ }
    \def \micst {micro\ste}
    \def \macst {macro\ste}
    \def \dof {degree of freedom\ }
    \def \dofs {degrees of freedom\ }
    \NewDocumentEnvironment{ lecture } {O{} O{} +b} 
	{\addtocounter{lecturecount}{1} % increase counter
		\phantomsection % invisible marker for hyperref
		\addcontentsline{toc}{subsection}{\textbf{Lecture~\thelecturecount:}~\textsl{#2}} % add to table of contents
		\begin{trivlist} % begin unmarked ("trivial") list
		\item \colorbox{LavenderBlush2}{\textbf{Lecture \thelecturecount.}~(\texttt{#1})~ \textit{#2}} % Header
		\item #3  % +b is for inputs that are long
		\end{trivlist} % end list
	}{}
\begin{document}
\maketitle
\input{../lectureformatting.tex}
    \begin{lecture}[Jan 05][States, Probability and Binomial Distribution]
        \begin{figure}[H]
            \centering
            \begin{tikzpicture}
                \draw[<->] (-2,0)--(2,0) node [anchor=west]{$x$};
                \draw[<->] (0,-2)--(0,2) node [anchor=south]{$p$};
                \draw[thick, black, fill] (1,1.5) circle(1pt);
            \end{tikzpicture}
            \caption{Phase space of 1-D \pa}
        \end{figure}
    \end{lecture}
    \begin{lecture}[Jan 10][Ensembles]
        Lagrange multipliers
        \begin{equation}
            S=-k\sum_rp_r\ln(p_r)
        \end{equation}
        Microcanonical ensemble: All accessible \micst are equally probable
    \end{lecture}
    \begin{lecture}[Jan 12][Finding total \micst]
        $N$ \pas in volume $V$ with energy between $E,E+\delta E$.
        Counting number of \micst by using phase space 
        \\
            simplifying example : a 1-D \pa has only $x$ and $p$. Plot in phase space
            Example in harmonic oscillator with ellipse and shading in phi(E) and Omega(E) Include text in caption explaining equations below it. 
        \\
        Moving to 3-D talk about \dofs and volume of $h_0$.
        \\
        Integrating to get Phi(E) with multiintegrals and then taylor approx to get Omega
        \\
        Quantum Description--> specify \micst with quantum numbers
        \\
        example with simp harmon oscill
    \end{lecture}
    \begin{lecture}[Jan 19][More on Microcanonical Ensemble]
        $\Omega(E)=\#$ of states with energy between $E+\delta E$
        \\
        Describing energy levels of each particle, think N-cube
        \\
        Now \pa can interact!! Mechanical interactions and thermal interactions(both macro descriptions). 
        \\
        1 isolated system at equilibrium$\rightarrow$ same system but with partition, now 2 systems. $A^0$ is comprised of $A,A'$. Macro parameters of $A^0$ are for both states (N,V,E,T,\ldots).
        \\
        \begin{description}
            \item [Thermal Interaction] External parameters of $A,A'$ are fixed but mean energy transferred from one system to the other as a result of purely thermal interactions called heat. Probabilities of energy states can change when systems interact $P(r)$
            \item [Mechanical Interaction] External Parameters of $A,A'$ change, one does \emph{work} on the other! This causes the mean energies of $A,A'$ to change. 
        \end{description}
        $$\overline{E}=\sum_rp_rE_r$$
        \\
        Pure thermal and purely mech example in inf sqwell
    \end{lecture}
\begin{lecture}[Jan 24]
    Pure thermal interaction changes $p_r$
    \\
    Pure mechanical interaction changes $E_r$
\end{lecture}
March 23rd
\\\\
Stable vs unstable equilibrium\\
use gibbs and helmholts free energy\\
isolated system vs in contact with temp or pressure reservoir\\
equilibrium in contact \\
system A changes states, excchanging heat or work. How much work can it do?\\
Isolated system\{Reservoir(A, T0,P0)--work source\}
A does $W^*$, energy, volume, and entropy increase by $\Delta E, \Delta V, \Delta S$ \\
change of volume and entropy is opposite of change in A bc of first law $W^*+\Delta E-\underbrace{P_0\Delta V_0+T\Delta S_0}_{reservoir}=0$
\\
2nd law says $\Delta(S+S_0)\geq0$ which means $\Delta S_0 \geq -\Delta S$\\
Sub in to first law equation \\
$W^*\leq -\Delta (E-T_0S+P_0V)=-\Delta G_0$
Not Gibbs free energy bc $T_0,P_0$ refer to reservoir, not to A, mixed with both variables.
$G_0=G$ when A is in equilibrium with res
\\
Since system is isolated, it starts and ends at $T,P=T_0,P_0$\\
meaning $W^*\leq -\Delta (E-TS+PV)=-\Delta G$ which is Gibb free energy. \\
System has fixed volume too, $\Delta V=0$\\
$W^*\leq -\Delta (E-T_0S)=-\Delta G$\\
if $T=T_0\rightarrow W^*\leq -\Delta F$ Helmholtz for A.\\
$W^*\leq -\Delta G$ in general\\
$W^*\leq -\Delta F$ for fixed volume\\
So $ -\Delta G$ and $-\Delta F$ are maximum work that can be done--> ``Free'' Energy!
\\
\\
Spontanetous process defined as process with $W^*=0$\\
$0\leq -\Delta G_0\rightarrow \Delta G_0\leq0\rightarrow (G_{0f}-G_{0i})\leq0\rightarrow \boxed{G_{0f}\leq G_{0i}}$
\\
For fixed volume: $\Delta F_0\leq0$\\
In equilibrium, systems in contact with T,P reservoirs, there can be no change in gibbs or helmholtz, meaning that final can't be less than initial, meaning that it must be a minimum! 
\\
analogue (?) with of $S=S_{max}$ for isolated sys in eq
\\
\\
Stability of thermo system. system A in reservoir, res has $T_0,P_0$. How do perterbations/fluctuations change $G_0$? Thinking abt stabel vs unstable equilibium. 
\\
Begin by characterizing A by $T,V$. 
\\
$$\pder{G_0}{T}[V]=0,\pder{G_0}{V}[T]=0$$
Expanding about minimum $G_min$.
$$G_0=G_{min}+\oo{2}\eval{\pder[2]{G_0}{T}[V]~}_{G=G_min}\qty(\Delta T)^2+\oo{2}\eval{\pder[2]{G_0}{V}[T]~}_{G=G_min}\qty(\Delta V)^2+\eval{\pder{G_0}{TV}~}_{G=G_min}\Delta T\Delta V+...$$
Now calculate derivs. 
$$\pder{G_0}{T}[V]=\pder{E}{T}[V]-T_0\pder{S}{T}[V]=T\pder{S}{T}[V]-T_0\pder{S}{T}[V]$$
So $T=T_0$
$$\pder{G_0}{V}[T]=\pder{E}{V}[T]-T_0\pder{S}{V}[T]=T\pder{S}{T}[V]-p-T_0\pder{S}{T}[V]+p_0=0$$
so $P=P_0$
$$\pder[2]{G_0}{V}[T]=T\pder[2]{S}{V}[T]-\pder{P}{V}[T]-T_0\pder[2]{S}{V}[T]=-\pder{P}{V}[T]=\oo{V\kappa}$$
$$\pder[2]{G_0}{T}[V]=T\pder[2]{S}{T}[V]+\pder{S}{T}[V]-T_0-\pder[2]{S}{T}[V]$$
$$=\pder{S}{T}[V]=\oo[C_V]{T}$$
$$\pder{G_0}{TV}=T\pder{S}{TV}-T_0\pder{S}{TV}=0$$
Combining them all together to get: 
$$G_0=G_{min}+\underbrace{\frac{C_V}{2T_0}\qty(\Delta T)^2+\oo{2V\kappa}\qty(\Delta V)^2}_{\text{ Must be $>0$ for stable}}$$
So stable if $C_V$and $\kappa >0$
\\\\
Monday March 28th\\
Phase Transitions as applicaitons of tdymic instability, aka $C_V$ or $\kappa <0$. Phase transitions accompany changes of internal energy of system.
\\
Change in crystaline structure.
\\
Change in conductor to superconductor.
\\\\
Must consider non ideal gasses. Van-der Waals gas. 
\\
For 1 mol of VDW gas: $$\qty(p+\frac{a}{v^2})(v-b)=RT$$
Consider isotherms on p-v diagram in Desmos, there is a section for low T where $\pder{P}{V}>0$ 
\\
System with 2 possible phases in contact with temp and pressure reservoir. Gibbs free energy. 
$$G(T,P)\equiv G(T,P)/\text{mole}\qquad \nu_i\equiv\text{\# moles in phase\ }i$$
For equilibrium $$G_T=\nu_1g_1+\nu_2g_2$$ must be a minimum. 
Constraint is that $$\nu_1+\nu_2=\nu$$
Three possibilities 
\begin{itemize}
    \item $g_1<g_2$ At equilibrum $G_T=\nu g_1$
    \item $g_1>g_2$ At equilibrium $G_T=\nu g_2$
    \item $g_1=g_2$ At equilibrium $G_T=\nu_1 g_1+\nu_2 g_2$
\end{itemize}
Phase diagram is a P-T diagram with lines that represent $g_i=g_j$ for two phases $i,j$
\\
For phase change:
\begin{gather*}
    g_1=g_2
    \\
    E_1-TS_1+PV_1=E_2-TS_2+PV_2
    \\
    (E_1-E_2)-\underbrace{T(S_1-S_2)}_{Q}+P(V_1-V_2)=0
\end{gather*}
First order phase transitions. \\
Characterized by: 
\begin{itemize}
    \item $g(T,P)$ is continuous across boundary between phases. 
    \item The first partial derivatives are discontinuous. $\pder{g}{P}[T]=v,\ \pder{g}{T}[P]=-s$ 
    So in first order latent heat is absorbed or given off. 
    take differentials a nd group together to get 
\end{itemize}
\begin{gather*}
    \Delta\pder{g}{P}=v_1-v_2=\Delta v\ \text{Change in molar volume}
    \\
    \Delta\pder{g}{T}=-(s_1-s_2)=-\Delta s\ \text{Change in molar entropy}
\end{gather*}
the amount of molar latent heat that must be absorbed is $l$ so $\Delta S=l/T$
$$\pder{P}{T}=\frac{\Delta S}{\Delta V}=\frac{l}{T\Delta V}$$
Clausius-Clapeyron equation. \\
First order phase transitions are characterized by latent heat produced. \\
Second order phase transitions: both g and 1st derivatives are cont but second derivatives are not,,, no latent heat. 
\\
Application of Clausius-Clapeyron equation. \\
consider liquid vapor, find vapor pressure in equilibrium with liquid. \\
Simplifying asssump: 
\begin{itemize}
    \item $\Delta V=V_{vap}-V_{liq}\approx V_{vap}\rightarrow V_{liq}<<V_{vap}$
    \item $PV=RT$
    \item Asssume $l\neq l(T)$
\end{itemize}
\begin{gather*}
    \der{P}{T}=\frac{l}{T}\oo{V_{vap}}=\frac{l}{T}\frac{P}{RT}=\frac{lP}{RT^2}
    \\
    \oo[dp]{p}=\frac{l}{RT^2}dt
    \\
    P=P_0e^{-l/RT}
\end{gather*}
\end{document}